

```{r} 
## Install Neotoma for R and other required packages ##
install.packages("devtools")
require(devtools)
install_github("neotoma", "ropensci")
require(neotoma)
install.packages("ggplot2")
```

```{r}
## Pull datasets from Neotoma ##
Ausdata <- get_site(loc = c(90,-80,179.9,0))
AusPolDatasets <- get_dataset(Ausdata)
AusPollenData <- get_download(dataset = AusPolDatasets)
```

Ind.Sites <- compile.downloads(AusPollenData)
counts<-rowSums(Ind.Sites[ , 11:ncol(Ind.Sites)], na.rm = TRUE)
Poa <- regexpr("Poaceae", colnames(Ind.Sites)) > 0
Poa.counts <- rowSums(Ind.Sites[ , Poa], na.rm=TRUE)

Ind.Sites["Poa.Counts"]<-Poa.counts2
Ind.Sites["Total.Counts"]<-counts
Poa.Prop<-(Ind.Sites[ ,221:222])

Poa.Prop <- data.frame(site = Ind.Sites$sitename, age = Ind.Sites$age, Pcount = Ind.Sites$Poa.Counts, Tcount = Ind.Sites$Total.Counts)
Poa.table<-na.omit(Poa.Prop)
```{r}
##split data table into ind. sites##
Blue<-Poa.table[Poa.table$site == "Blue Lake", ]
George<-Poa.table[Poa.table$site == "Lake George, "]
Yarl<-Poa.table[Poa.table$site == "Yarlington Tier", ]
```

```{r}
##Set up Blue Lake for model run
Blue$age<- -Blue$age
Blue<-Blue[order(Blue$age), ]
nT<-nrow(Blue)
plot(Blue$age, Blue$Pcount/Blue$Tcount)
lines(Blue$age, Blue$Pcount/Blue$Tcount)

#Set model params

y<-Blue$Pcount
n<-Blue$Tcount


```

##Bugs Code
hmm<- function() {
  mu ~ dunif(-5,5)
  rho ~ dunif(-1,1)
  tau ~ dunif(0,100)
  logitTheta ~ dnorm(0,1e-06)
  tau2Inv<- 1/(tau^2)
  
  logitTheta[1] ~ dnorm(mu + rho * (logitTheta - mu), tau2Inv)
    theta[1]<- exp(logitTheta[1])/(1+exp(logitTheta[1]))
      y[1] ~ dbin(theta[1], n[1])
        for (i in 2:nT) {
        logitTheta[i] ~ dnorm(mu+rho*(logitTheta[i-1]-mu), tau2Inv)
        theta[i] <- exp(logitTheta[i])/(1+exp(logitTheta[i]))
          y[i] ~ dbin(theta[i], n[i])
        }
}

#End of Bugs code

```{r}
##fit the model through R
# increase num of interations!
out <- jags(data = list(nT = nT, n = n, y = y), parameters.to.save = c("rho", 
    "tau", "theta"), n.chains = 1, n.iter = 10000, n.burnin = 2000, model.file = hmm, 
    DIC = FALSE)
```

```{r}

## construct resulting mcmc
out.mcmc <- as.mcmc(out)[[1]]
thetaHat <- Blue$Pcount/Blue$Tcount

colNames <- dimnames(out.mcmc)[[2]]
whichTheta <- grep("theta", colNames)
thetaPost <- out.mcmc[201:1000, whichTheta]
thetaNames <- dimnames(thetaPost)[[2]]
index <- gsub("theta\\[", "", thetaNames)
index <- as.numeric(gsub("\\]", "", index))
thetaPost <- thetaPost[, order(index)]

##create output figures
par(mfrow = c(1, 1))
plot(Blue$age, thetaHat, col = "red")
quants <- apply(thetaPost, 2, quantile, c(0.025, 0.975))
polygon(cbind(c(Blue$age, rev(Blue$age), Blue$age[1]), c(quants[1, ], rev(quants[2, ]), quants[1, 1])), border = NA, col = "lightblue")
lines(Blue$age, thetaHat, col = "red")
points(Blue$age, thetaHat, col = "red")
lines(Blue$age, colMeans(thetaPost), col = "blue")
title("Grass Presence at Blue Lake, New South Wales")

library("ggmap"")
library("mapproj")
map <- get_map(location = 'Australia', zoom = 4)
ggmap(map)
