
```{r}
## Install Neotoma for R and other required packages ##
install.packages("devtools")
require(devtools)
install_github("neotoma", "ropensci")
require(neotoma)
require(R2jags)
```
```{r}
## Pull datasets from Neotoma ##
Ausdata <- get_site(loc = c(90,-80,179.9,0))
AusPolDatasets <- get_dataset(Ausdata)
AusPollenData <- get_download(dataset = AusPolDatasets)
```
Ind.Sites <- compile_downloads(AusPollenData)
counts<-rowSums(Ind.Sites[ , 11:ncol(Ind.Sites)], na.rm = TRUE)
Poa <- regexpr("Poaceae", colnames(Ind.Sites)) > 0
Poa.counts <- rowSums(Ind.Sites[ , Poa], na.rm=TRUE)
Poa.counts2<-as.data.frame(Poa.counts)
Ind.Sites["Poa.Counts"]<-Poa.counts2
Ind.Sites["Total.Counts"]<-counts
Poa.Prop<-(Ind.Sites[ ,221:222])
Poa.Prop <- data.frame(site = Ind.Sites$sitename, age = Ind.Sites$age, Pcount = Ind.Sites$Poa.Counts, Tcount = Ind.Sites$Total.Counts)
Poa.table<-na.omit(Poa.Prop)
```{r}
##split data table into ind. sites##
Blue<-Poa.table[Poa.table$site == "Blue Lake", ]
George<-Poa.table[Poa.table$site == "Lake George", ]
Yarl<-Poa.table[Poa.table$site == "Yarlington Tier", ]
```
```{r}
##Set up Blue Lake for model run
Blue$age<- -Blue$age
Blue<-Blue[order(Blue$age), ]
nT<-nrow(Blue)
plot(Blue$age, Blue$Pcount/Blue$Tcount)
lines(Blue$age, Blue$Pcount/Blue$Tcount)
#Set model params
y<-as.integer(Blue$Pcount)
n<-as.integer(Blue$Tcount)
```
##Bugs Code
hmm <- function(){
# mu between -5,5 allows theta to go between 0 and 1. Good 'approx' values
mu ~ dunif(-5, 5)
rho ~ dunif(-1, 1)
tau ~ dunif(0, 100)
logitTheta0 ~ dnorm(0, .000001)
tau2Inv <- 1/(tau^2)
# latent process evolution and likelihood
logitTheta[1] ~ dnorm(mu + rho*(logitTheta0 - mu), tau2Inv)
theta[1] <- exp(logitTheta[1])/(1+exp(logitTheta[1]))
y[1] ~ dbin(theta[1], n[1])
for(i in 2:nT){
logitTheta[i] ~ dnorm(mu + rho*(logitTheta[i-1] - mu), tau2Inv)
theta[i] <- exp(logitTheta[i])/(1+exp(logitTheta[i]))
y[i] ~ dbin(theta[i], n[i])
}
}#hmm()
#End of Bugs code
```{r}
##fit the model through R
# increase num of interations!
out <- jags(data = list(nT = nT, n = n, y = y),
parameters.to.save = c("rho", "tau", "theta"), n.chains = 1,
n.iter = 10000, n.burnin = 2000, model.file = hmm, DIC = FALSE)
```
```{r}
## construct resulting mcmc
out.mcmc <- as.mcmc(out)[[1]]
thetaHat <- Blue$Pcount/Blue$Tcount
colNames <- dimnames(out.mcmc)[[2]]
whichTheta <- grep("theta", colNames)
thetaPost <- out.mcmc[201:1000, whichTheta]
thetaNames <- dimnames(thetaPost)[[2]]
index <- gsub("theta\\[", "", thetaNames)
index <- as.numeric(gsub("\\]", "", index))
thetaPost <- thetaPost[, order(index)]
##create output figures
par(mfrow = c(1, 1))
plot(Blue$age, thetaHat, col = "red")
quants <- apply(thetaPost, 2, quantile, c(0.025, 0.975))
polygon(cbind(c(Blue$age, rev(Blue$age), Blue$age[1]), c(quants[1, ], rev(quants[2, ]), quants[1, 1])), border = NA, col = "lightblue")
lines(Blue$age, thetaHat, col = "red")
points(Blue$age, thetaHat, col = "red")
lines(Blue$age, colMeans(thetaPost), col = "blue")
title("Grass Presence at Blue Lake, New South Wales")
out.mcmc <- as.mcmc(out)[[1]]


```
```{r}
##Repeat setup for Lake George
George$age<- -George$age
George<-George[order(George$age), ]
nT<-nrow(George)
plot(George$age, George$Pcount/George$Tcount)
lines(George$age, George$Pcount/George$Tcount)
y<-as.integer(George$Pcount)
n<-as.integer(George$Tcount)
```
```{r}
#Run Model again for Lake George
##Bugs Code
hmm <- function(){
# mu between -5,5 allows theta to go between 0 and 1. Good 'approx' values
mu ~ dunif(-5, 5)
rho ~ dunif(-1, 1)
tau ~ dunif(0, 100)
logitTheta0 ~ dnorm(0, .000001)
tau2Inv <- 1/(tau^2)
# latent process evolution and likelihood
logitTheta[1] ~ dnorm(mu + rho*(logitTheta0 - mu), tau2Inv)
theta[1] <- exp(logitTheta[1])/(1+exp(logitTheta[1]))
y[1] ~ dbin(theta[1], n[1])
for(i in 2:nT){
logitTheta[i] ~ dnorm(mu + rho*(logitTheta[i-1] - mu), tau2Inv)
theta[i] <- exp(logitTheta[i])/(1+exp(logitTheta[i]))
y[i] ~ dbin(theta[i], n[i])
}
}#hmm()
require(R2jags)
out <- jags(data = list(nT = nT, n = n, y = y),
parameters.to.save = c("rho", "tau", "theta"), n.chains = 1,
n.iter = 10000, n.burnin = 2000, model.file = hmm, DIC = FALSE)
```
```{r}
#Output for Lake George
## construct resulting mcmc
out.mcmc <- as.mcmc(out)[[1]]
thetaHat <- George$Pcount/George$Tcount
colNames <- dimnames(out.mcmc)[[2]]
whichTheta <- grep("theta", colNames)
thetaPost <- out.mcmc[201:1000, whichTheta]
thetaNames <- dimnames(thetaPost)[[2]]
index <- gsub("theta\\[", "", thetaNames)
index <- as.numeric(gsub("\\]", "", index))
thetaPost <- thetaPost[, order(index)]
##create output figures
par(mfrow = c(1, 1))
plot(George$age, thetaHat, col = "red")
quants <- apply(thetaPost, 2, quantile, c(0.025, 0.975))
polygon(cbind(c(George$age, rev(George$age), George$age[1]), c(quants[1, ], rev(quants[2, ]), quants[1, 1])), border = NA, col = "lightblue")
lines(George$age, thetaHat, col = "red")
points(George$age, thetaHat, col = "red")
lines(George$age, colMeans(thetaPost), col = "blue")
title("Grass Presence at Lake George, New South Wales")
out.mcmc <- as.mcmc(out)[[1]]
plot(out.mcmc)
```
```{r}
##Repeat setup for Yarlington Tier
Yarl$age<- -Yarl$age
Yarl<-Yarl[order(Yarl$age), ]
nT<-nrow(Yarl)
plot(Yarl$age, Yarl$Pcount/Yarl$Tcount)
lines(Yarl$age, Yarl$Pcount/Yarl$Tcount)
y<-as.integer(Yarl$Pcount)
n<-as.integer(Yarl$Tcount)
```
```{r}
#Run Model again for Yarlington Tier
##Bugs Code
hmm <- function(){
# mu between -5,5 allows theta to go between 0 and 1. Good 'approx' values
mu ~ dunif(-5, 5)
rho ~ dunif(-1, 1)
tau ~ dunif(0, 100)
logitTheta0 ~ dnorm(0, .000001)
tau2Inv <- 1/(tau^2)
# latent process evolution and likelihood
logitTheta[1] ~ dnorm(mu + rho*(logitTheta0 - mu), tau2Inv)
theta[1] <- exp(logitTheta[1])/(1+exp(logitTheta[1]))
y[1] ~ dbin(theta[1], n[1])
for(i in 2:nT){
logitTheta[i] ~ dnorm(mu + rho*(logitTheta[i-1] - mu), tau2Inv)
theta[i] <- exp(logitTheta[i])/(1+exp(logitTheta[i]))
y[i] ~ dbin(theta[i], n[i])
}
}#hmm()
require(R2jags)
out <- jags(data = list(nT = nT, n = n, y = y),
parameters.to.save = c("rho", "tau", "theta"), n.chains = 1,
n.iter = 10000, n.burnin = 2000, model.file = hmm, DIC = FALSE)
```
```{r}
#Output for Yarlington Tier
## construct resulting mcmc
out.mcmc <- as.mcmc(out)[[1]]
thetaHat <- Yarl$Pcount/Yarl$Tcount
colNames <- dimnames(out.mcmc)[[2]]
whichTheta <- grep("theta", colNames)
thetaPost <- out.mcmc[201:1000, whichTheta]
thetaNames <- dimnames(thetaPost)[[2]]
index <- gsub("theta\\[", "", thetaNames)
index <- as.numeric(gsub("\\]", "", index))
thetaPost <- thetaPost[, order(index)]
##create output figures
par(mfrow = c(1, 1))
plot(Yarl$age, thetaHat, col = "red")
quants <- apply(thetaPost, 2, quantile, c(0.025, 0.975))
polygon(cbind(c(Yarl$age, rev(Yarl$age), Yarl$age[1]), c(quants[1, ], rev(quants[2, ]), quants[1, 1])), border = NA, col = "lightblue")
lines(Yarl$age, thetaHat, col = "red")
points(Yarl$age, thetaHat, col = "red")
lines(Yarl$age, colMeans(thetaPost), col = "blue")
title("Grass Presence at Yarlington Tier, Tasmania")
out.mcmc <- as.mcmc(out)[[1]]
plot(out.mcmc)